---
title: "Machine Learning"
author: 
  - Bach, Helena ^[Università di Bologna, xxx]
  - Fleming, María^[Università di Bologna, xxx]
  - Karampelas, Petro^[Università di Bologna, xxx]
  - Romero, Pablo José^[Universidad Nacional de Córdoba, xxx]
date: "30/12/2021"
output:
  pdf_document:
    extra_dependencies: ["float"]
    number_sections: False
header-includes:
  \usepackage{floatrow}
  \floatsetup[figure]{capposition=top}
  \usepackage[font={large}]{caption}
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H", out.extra = "")
```

# Introduction

|   In this document the following R packages are made use of:

```{r packages, warning=FALSE, message=FALSE}

library(corrplot)
library(caTools)
library(biotools)
library(ggplot2)
library(MASS)
library(xtable)
library(dplyr)
library(tidyr)
library(caret)
library(kableExtra)

```

# Descriptive Analysis

|   The data set used in this project collects information about a Portuguese bank marketing campaign aiming to get customers to subscribe to a term deposit. The data set contains 45.211 observations and 17 different variables: 11 categorical and 6 numerical.
|   Firs of all, the data will be uploaded, the variables classified, the train and test sample created and a exploratory analysis will be done.

```{r exploratoryAnalysis}

bank <- read.csv(file="bank-full.csv", header=TRUE, sep=";")
summary(bank)
str(bank)

#variables that should be treated as categorical

bank$marital <- as.factor(bank$marital)
bank$education <- as.factor(bank$education)
bank$default <- as.factor(bank$default)
bank$housing <- as.factor(bank$housing)
bank$job <- as.factor(bank$job)
bank$y <- as.factor(bank$y)

#create test and training sample

set.seed(101)
sample = sample.split(bank[,1], SplitRatio = .75)
train <- bank[sample, ]
test <- bank[-sample, ]

```

Figure \ref{fig:boxplot} shows the box plot of the numerical variables. It can be noticed that \emph{age} follows a very similar distribution for clients who subscribed to a term deposit and for those who did not, indicating that the age of the customer won't play a significant role when predicting the behavior of clients. On the other hand, from the box-plot of \emph{duration} it can be inferred that high duration (long contact duration in second) might have a positive impact on the subscription rate. 

```{r boxplot, dev='png', fig.cap="\\label{fig:boxplot}Boxplot", fig.height=8}
#########DESCRIPTIVE ANALYSIS#########

par(mfrow=c(3,2))
boxplot(age ~ y, data = bank, main="Age")
boxplot(balance ~ y, data = bank, main="Balance")
boxplot(day ~ y, data = bank, main="Day")
boxplot(duration ~ y, data = bank, main="Duration")
boxplot(pdays ~ y, data = bank, main="Pdays")
boxplot(previous ~ y, data = bank, main="Previous")

```

# Logit

|   The logistic regression models the probability of belonging to one class, under the assumption that the dependent variable follows a binomial distribution. It is a classification technique that allows to predict a dichotomous variable. 

Given $X$ (explanatory variables) we can represent the probability  that the client has subscribed to a term deposit (Y) as $p(X)=P(y=yes|X)$, using the logistic function to ensure that the output lies between 0 and 1.

\begin{center}
$p(X)=\frac{e^{\beta_0+\beta_1 \cdot X}}{1+e^{\beta_0 + \beta_1\cdot X}}$
\end{center}

```{r logitmod}

mylogit <-  glm(y ~., data = train,
                family = "binomial")

tabl <- summary(mylogit)

kable(tabl$coefficients,
      caption = "\\label{fig:logit}Logit summary",
      format ="latex", 
      align = "c",
      table.envir = "figure")

```


From Table \ref{fig:logit} (included in the annex) it can be inferred which variables are significant when determining the outcome of the term deposit subscription. Working as a housemaid, entrepreneur, manual labor or being self-employed has a significant negative effect on the probability of subscribing to the term deposit, while being retired or a student increases the probability. In terms of education, completing  a secondary or tertiary level of education, increases the probability of subscribing. Clients who are married, have a housing or personal loan are less likely to subscribe to the term deposit. Accounts with  a large average yearly balance have higher odds of accepting the marketing offer. In line with Figure \ref{fig:boxplot}, duration has a positive significant effect, so clients that were contacted for a longer time (seconds) are more likely to subscribe.

To evaluate the performance of the logistic model defined previously, we create a confusion matrix. The confusion matrix compares the predicted values obtained from the model to the actual values.

```{r prediclogit}

probabilities <- mylogit %>% predict(test, type = "response")
test$predicted <- ifelse(probabilities > 0.5, 'yes', 'no')

#Confusion matrix
cmatrix <- confusionMatrix(table(test$y, test$predicted))
                           
kable(cmatrix$table,
      caption = "\\label{fig:confmat}Confusion Matrix",
      format ="latex", 
      align = "c",
      table.envir = "figure")
```
